### 2018-09-14 revision/responseToreviewers.txt
add txt version fo response

As a preamble, we would like to thank the editor and the reviewer for
their comments and suggestions. Following these comments, we made
several changes to the article, which are summarized here. The next
sections list our answers to each of the reviewer’s comments, with
references to the revised manuscript (page, column, and paragraph) where
appropriate.

Answers to Reviewer 1
=====================

Contribution and state of the art
---------------------------------

1.  *The authors miss some important references when they review the state of the art related to the addressed problem. For instance, those articles published within the DCASE 2017 challenge, which are not included in the paper. Moreover, other specific relevant works should be referenced, such as:
\begin{itemize}
\item Onur Dikmen and Annamaria Mesaros. "Sound event detection using non-negative dictionaries learned from annotated overlapping events", 2013 IEEE Workshop on Applications of Signal Processing to Audio and Acoustics (WASPAA), where also a threshold-based NMF approach is applied to a very similar problem.\\
And the work previously published by the same authors in DCASE 2016:
\item Jean-Rémy Gloaguen, Arnaud Can, Mathieu Lagrange, Jean-François Petiot. "Estimating traffic noise levels using acoustic monitoring: a preliminary study." Detection and Classification of Acoustic Scenes and Events (DCASE 2016), September 2016, Budapest, Hungary.
\end{itemize}
Both papers should be included as related work explicitly explaining the main differences with respect to the presented work.*

    The introduction has been enriched with the two
    proposed references. The differences between the present
    contribution and the work of Dikmen & al @dikmen2013sound and our
    previous work @gloaguen2016estimating is now clearly stated. We also
    added the work of Mesaros & al @mesaros_sound_2015

2.  *The authors argue that “Prior to data assimiliation, the issue of
    the correct estimation of the traffic sound level from acoustic
    measurements is still unsolved \[31\]”, being \[31\] a reference
    from 2011. The authors should rephrase the sentence or change the
    reference in order to update the current state of the art in
    this field. For instance, only reference \[35\] is included
    describing deep learning approaches, which is one of the relevant
    approaches in this research field during the last years.*

     The sentence has been rephrased with updated
    references (please see Section 1) :’Prior to data assimilation, the
    issue of the correct estimation of the traffic sound level from
    acoustic measurements begin to be studied
    @leiba2017large@socoro_anomalous_2017. As the urban sound
    environment is a complex environment gathering lots of different
    sounds (car passages, voices, whistling bird, car horn, airplanes…)
    that overlap, the traffic sound level estimation based on
    measurements is not a trivial task.’

3.  *As a final suggestion, I would recommend the authors completing the
    review of the state of the art by including a specific section about
    the Related Work in the manuscript. Although this section is not
    mandatory according to the guide for authors in this journal, it
    would help to clarify the contribution of the authors with respect
    to previous similar approaches.*

     This section has been added as Section 2 of the
    updated manuscript where related work is now presented and where we
    highlight the differences between those works and the approach taken
    in our contribution.

Metric and results
------------------

1.  *Since Lp values are in dB, you cannot average them linearly as
    considered in your proposed metric MAE of eq. (15). The values must
    be averaged in their linear representation before converting them
    into dBs again. This error may have a dramatic impact on
    your results.*\

    The use of arithmetic operators on logarithmic values is a common
    practice in the urban acoustics field when dealing with errors (see
    Aumond et al. “Kriging-based spatial interpolation from measurements
    for sound level mapping in urban areas.” @aumond2018kriging and
    Morillas et al. “Uncertainty evaluation of continuous
    noise sampling.” @morillas2014uncertainty). In order to better
    position our design choice in terms of metrics a paragraph is now
    added in Section 4.2.3.

    We acknowledge that strictly physically speaking this might be
    considered as a mistake but we chose to follow this approach as we
    believe it allows the calibration of the traffic sound level and to
    monitor metrics that are less sensitive to high values. As the
    errors are treated in a statistical point of view we consider that
    this operation is well suited.

2.  *First, the results obtained when considering fc=20Khz (“fc = 20 kHz
    is equivalent to consider all the sound mixtures without distinction
    between traffic and others sound sources”) show the lowest MAE for
    TIR=<span>6,12</span>, where “the traffic is the main sound source”.
    –&gt; Have the authors realized the consequences of these results?
    From these them someone can conclude that it is better to process
    nothing when interfering sources are 6 and 12 dB lower with respect
    to the traffic noise level!!*\
     This results suggests that the user does not apply any
    treatment in cases where road traffic predominates. This is not
    inconsistent with the purpose of this paper for two reasons: (i) a
    user would not call for signal processing in situations where road
    traffic is largely dominating such as in the vicinity of a ring
    road, (ii) the TIR value is an unknown parameter: the high errors of
    the filter for negative TIR values discredit its use. We made the
    precision : ’Even if the performances for the positive $TIR$ values
    are high, it has to be reminded that in practice, the $TIR$ value is
    not known. In consequence, without any prior knowledge, this
    approach cannot be applied.’ (please see the beginning of
    Section 5).

3.  *Second, for the remaining studied scenarios, i.e.,
    TIR=<span>-12,-6,0</span>, the lowest averaged MAE values are
    obtained by the Sem-NMF, and not by the proposed TI-NMF. Therefore,
    from these results, it cannot be concluded that TI-NMF is the best
    approach in the considered experimental framework. Even more, if you
    remove from Table 4 the aforementioned values obtained with
    TIR=<span>-12,-6</span>, Sem-NMF becomes the best approach with an
    averaged MAE of 2.12, whereas TI-NMF yield an averaged MAE of 3.14,
    following the same analysis scheme proposed by the authors.*

     We made a precision at the end of section 2 : ’The aim
    is to find the type of NMF that gives the minimal error on the
    reconstruction of the traffic noise signal on the whole corpus to be
    applied to all the different possible urban noise environments.’

    Furthermore, as now explained at the end of section 4.1, in urban
    sound environments, TIRs are mainly focused in the range
    $\left[-6, 12\right]$ dB @gloaguen_creation_2017. $TIR$ = -12 dB is
    a rarer case in urban environment but it is still added in order to
    show the limits of the evaluated methods.

4.  *Finally, I would recommend the authors to consider including a
    statistical analysis to prove that differences between the different
    approaches are statistically significant or not.*

     As suggested by the reviewers, we added a Student
    t-test in Section 5 with the results exposed in Table 3. This test
    revealed that the distribution of the traffic sound level
    estimations between the best methods exposed in Table 2 are
    statistically different.

5.  *Last but not least, please review the results of the low pass
    filter approach with fc=20kHz in Table 4 and 3, since the global
    averaged MAE in Table is 4.54, while in Table 3 is 4.69.*

     The value at TIR = 6 was a misprint, it has
    been corrected.

Meaningfulness of the experiments
---------------------------------

1.  *The baseline approach considered by the authors is a low-pass
    filter with different cut-off frequencies “fc=<span>500, 1k, 2k, 5k,
    10k, 20k</span> Hz.”. First, this baseline may be understood as too
    simple for the problem at hand by the research community since, at
    least, supervised NMF has been already considered in the literature.
    Second, no comparison with \[38\] or other previous attempts to
    apply NMF to the problem at hand is provided. To this aim, it would
    be especially interesting the comparison with (Dikmen and
    Mesaros, 2013). Finally, after introducing the fc sweep in section
    3.2, only fc=<span>500, 20k</span>Hz are the subsequently considered
    with no further justification.*

     The baseline of the low-pass filter, even if it seems
    too simple, is very effective in some cases. As in the environmental
    acoustics community, this method would have been easily seen as a
    tool, we decided to consider it. We summarized all the results for
    the low pass filter according to the cut-off frequency in order to
    be more complete (Figure 9). As for the comparison with (Dikmen and
    Mesaros, 2013), even if they use the NMF framework, they develop a
    detection tool that is not adapted to this study. We believe that it
    was wiser to use a low-pass filter as a low complexity baseline.

2.  *The authors have considered a database composed of artificially
    mixed sound sources. They argue this decision as follows: “The use
    of simulated sound scenes is mandatory for rigorous experimental
    validation as it offers a high level of control on the design of the
    scenes” –$>$ I would suggest removing the term “mandatory” from
    the argumentation. Maybe the sentence could be rewritten as: “The
    use of simulated sound scenes allows…”*

     The sentence have been rephrased to : ’The use of
    simulated sound scenes allows rigorous experimental validation …’
    (section 2).

3.  *Moreover, for this reviewer it is not clear what is considered
    traffic and not traffic by the authors in order to compute
    $Lp_traffic$. The “environmental sound scene corpus” is composed of
    “constant traffic noise” and “passing cars”, which have been
    recorded in a real life environment by the authors. The car pass-bys
    are then used in several figures. For instance, Fig.4 shows an
    example created with the SimScene software showing background road
    traffic noise together with three overlapped sound sources: car
    horn, car passage and whistling birds. Does this mean that the car
    pass-by is considered as an interfering noise source? I would
    recommend the authors to clarify this point since a car pass-by is
    also traffic noise.*

     This important point is now clarified in the
    manuscript: ’What is called traffic component is the sum of the road
    traffic background noise and the sound events generated by the
    passing car class. On the contrary, the *interfering* sound class
    includes all the other sound sources not related to it. Car horn
    sound class belongs to this component as it is considered as a
    warning signal’ (Section 4.1).

4.  *Furthermore, the authors provide a lot of information about the
    configurations of the different approaches. However, I would
    recommend including specific information about the total duration of
    the train and test databases (e.g., not only saying “750 sound
    mixtures”), besides explicitly explaining the test scheme considered
    to guarantee that train and test data is decoupled when evaluating
    the considered sub-classes.*

     Those numbers are now added to the manuscript. For the
    test database, see section 4.2.1: ’each scene during 30 seconds, it
    leads to a full duration of 6 hours and 30 minutes.’ and for the
    train database, see section 4.1: ’The train database is composed of
    53 audio files of isolated passing cars with a 18 minutes cumulative
    duration.’

5.  *Finally, it is worth mentioning that considering
    TIR=<span>-6,-12,0,6,12</span> may yield to conclusions far from
    what can be found in real-life environments, as discussed in the
    literature (e.g. see \[42\]). Thus, I would recommend rephrasing the
    following sentence: “This database allows the creation of a wide
    diversity of realistic urban sound scenes from the road traffic
    point of view \[16\]” accordingly*

     The sentence is now rephrased in section 4.1 to: ’This
    database allows the creation of a wide diversity of urban sound
    scenes from the road traffic point of view’ …In addition, at the end
    of section 4.1 the motivation behind the choice of the range of TIR
    values is now discussed in depth, highlighting the fact that,
    indeed, the range is definitely wider than what can be found in
    realistic settings. The results are also now discussed by
    considering this matter into account.

Minors comments
---------------

1.  *In the introduction, the authors explain that road traffic noise
    maps show Lden and LN considering the A-weighted equivalent sound
    pressure level, thus, expressing the noise levels in dBA. However,
    their work only focuses on the equivalent sound pressure level in
    dB, why?*

     This choice is now explained in section 4.2.2: ’The
    $A$-weighting of the sound levels is not considered here as it
    decreases the low frequencies levels where the road traffic
    components are mainly present.’

2.  *Sections 2.1, 2.2 and 2.3 describing NMF, Supervised and
    Semi-Supervised NMF could be removed after extending the
    related work. This way, the manuscript could emphasize the proposal
    TI-NMF with respect to similar works.*

     Our contribution is targeted in part to the urban
    acoustic community. As NMF is not a well-known tool in this
    community, we thus prefer to keep the presentation of the
    fundamentals of this method.

3.  *The authors state that “Sup-NMF and Sem-NMF updates are computed
    for 400 iterations, which is sufficient to reach convergence. TI-NMF
    is performed on a lower number of iterations (60) to prevent W to
    not deviate too much from the initial dictionary.” –&gt; Please,
    justify this qualitative argumentation.*

     We now consider 100 as the number of iterations for
    each method, see Section 4.2.2: ’100 iterations are performed for
    every NMF types’.

4.  *Different kind of space reduction and data representation are
    included to reduce the computational time, e.g. K-means and third
    octave bands. –&gt; I would recommend the authors to include some
    kind of diagram to better explain their approach, and, if possible,
    to include some experiments to argue their argumentation, for
    instance, to explain why you have chosen third octave bands in front
    of other possible parameterization such as MFCC.*

     About the choice of third-octave bands, we specify in
    section 4.2.2 ’The spectrogram $\mathbf{V}$ and the dictionary
    $\mathbf{W}$ are expressed with third octave bands ($F$ = 29). This
    coarser method allows us to reduce the dimensionality and then
    decrease the computation time. Furthermore, by expressing the
    frequency axis on a log frequency axis, the low frequencies, where
    the traffic energy is focused, are described more finely than the
    high frequencies. Experimental validation consistently showed that
    considering third octave bands do not impact the performance of the
    estimator studied in this paper. But, most of all, it is a suited
    representation to this sound environment as this kind of
    representation is widely used in the urban acoustic field, compare
    to MFCC for instance.’. Also, we add a block diagram which details
    the dictionary building and makes clearer these different steps
    (please see Figure 7).

5.  *“The threshold t is set between 0.30 and 0.70 with a step of 0.01”
    – &gt; Please include some argumentation or reference to justify the
    range selected for the experiments beyond the ones reported in
    Table 3. Please, include more information of these experiments.*

     This matter is now discussed in Section 4.2.2. ’For
    TI-NMF, a preliminary study showed that the range of threshold
    values can be evaluated between 0.30 and 0.70. An increment step of
    0.01 has been considered as being sufficiently precise.’

6.  *All the analyses are done considering t=0.42 as the best threshold
    for the TI-NMF approach. Nevertheless, the authors stand that “With
    a low threshold, it is possible to decrease the error”, and present
    some examples subsequently to demonstrate that for high TIR=12 dB,
    the averaged MAE can be decreased by reducing the threshold value
    from t=0.42 to t=0.30. Correspondingly, the threshold should be
    moved from t=0.42 to t=0.55 to improve the results for low
    TIR=-12 dB. To what extent these values show better performance or
    they are due to overfitting?*

     This improvement is due to a decrease of the number of
    elements considered as traffic components. As in TIR = -12 dB, the
    interfering class is more present, it takes more basis in W’. By
    increasing the threshold, we put aside more elements from W’ that
    can belong to this sound class. It is equivalent, in detection task,
    to a reduction of the number of false-positive. It is now discussed
    further in the end of the section 5: ’It would be possible to
    decrease the error by adapting the threshold according to the $TIR$.
    For instance at $TIR$ = 12 dB, the average error on all the
    sub-classes would decrease to 0.30 ($\pm$ 0.11) with $t$ = 0.34
    where 191 elements are considered as traffic component. In the
    contrary, at $TIR$ = -12 dB, by increasing the threshold to 0.53,
    the error would decrease to 4.49 ($\pm$ 1.75) with in average 73
    elements considered as *traffic* elements.’

7.  *I would recommend including all the tuning and sweep analysis
    within an Appendix at the end of the manuscript.*

     As lots of results have been generated by those
    experiments, we chose to detail only some of the parameters that may
    influence the performance of the studied algorithms.

Some further details
--------------------

1.  *A -12 is missing in the enumeration of TIR values after eq. (12).
    Nevertheless, I would recommend removing this enumeration from the
    explanation of the equation since these values should be included in
    the Experiments section, where they are indeed included.*

     This part has been removed.

2.  *Tables 1 and 2 could be fused within the same Table.*

     The Tables have been fused.

3.  *In Section 3.2.2, wt <span>0.5 1</span> –&gt; a comma is missing
    between 0.5 and 1, i.e., <span>0.5, 1</span>*

     The misprint is now corrected.

4.  *Fig 8. The y-axis range does not show the minimum values of the
    alarm Lp,1s lower than 50 dB. Please, include them to better
    understand the impact of the alarm on the Lp computation. Moreover,
    to better understand the example, I would recommend the authors to
    append the spectrogram of the example to the Lp.1s (dB) figure.*

     Outside these emergences, the car horn sound level is
    null, that is why we limited the $y$-axis range to \[45 85\] dB to
    see, at best, the curves. In addition, we followed the reviewer’s
    recommendation by adding the spectrogram of the audio scene where
    the car horn sounding is visible (please see Figure 11).

5.  *Fig 9. Caption: “according to the the best” –$>$ remove one ’the*

     The caption title has been fixed.

English writing
---------------

1.  *Section 1: “Prior to data assimiliation,” –$>$ assimilation*

     The typo is corrected.

2.  *Section 2.4: “Usually in unsupervised,” a name is missing in
    the sentence.*

     The sentence is now: ’Usually in unsupervised
    learning,…’

3.  *Section 4: “First, the errors produced by the filter are detailed”
    –&gt; the errors are not produce by the filter itself. Please,
    rephrase the sentence to better explain the readers what are
    referring to.*

     The sentence is now rephrased to: ’First, the errors
    calculated from the traffic sound estimation of the low-pass filter
    are detailed’

4.  *“the traffic elements selected in TI-NMF do not activated when”
    –&gt; are not activated?*

     The sentence is now rephrased to: ’the traffic
    elements selected in TI-NMF are not activated’

5.  *Section 5. “In this work the non negative matrix factorization
    framework was used” –&gt; I would recommend using present perfect
    instead of past tense in the conclusions.*

     The sentence is now rephrased to: ’In this work the
    non negative matrix factorization framework is used …’

Answers to Reviewer 2
=====================

1.  *As the previous work has been already published, the authors should
    provide references to it and clearly state the new contributions of
    the presented paper. I believe there may be a substantial overlap
    with the authors’ paper Gloaguena, J. R., Cana, A., Lagrangeb, M.,
    $\&$ Petiotb, J. F. Estimation du niveau sonore du trafic routier au
    sein de mixtures sonores urbaines par la Factorisation en Matrices
    Non négatives, 2018.*

     The paper ’Gloaguen, J. R., Can, A., Lagrange, M.,
    $\&$ Petiot, J. F. Estimation du niveau sonore du trafic routier au
    sein de mixtures sonores urbaines par la Factorisation en Matrices
    Non négatives, 2018’ is a congress article without review submitted
    after the submission of the present paper. It deals with a similar
    approach but on a different sound corpus. Furthermore, as it is
    written in french we chose to alternatively cite ’Gloaguen, J.R.,
    Can, A., Lagrange, M., Petiot, J.F.: Estimating traffic noise levels
    using acoustic monitoring: A preliminary study’ which is written in
    English and deals with the same corpus.

2.  *Some sentences in the abstract are not well formulated and are hard
    to follow. E.g. “The task being to the best of our knowledge never
    been considered in the literature, we propose an experimental
    protocol to validate the studied approaches that complies with
    standard reproducible research recommendations.” should be splitted,
    “being” should be omitted; “raise this kind of innovative
    approaches” - the verb “raise” is confusing; “based on non-negative
    factorization framework” - the word “matrix” is missing. The authors
    may want to be more specific about the obtained results.*

     The abstract have been rewritten to take those
    requests into account: ’The task is to the best of our knowledge
    never been considered in the literature. We propose…’, ’Among the
    technical issues that bring this kind of innovative approaches,…’,
    ’several techniques based on the non-negative matrix
    factorization framework’.

3.  *“Their spectra can be obtained and be a basis of W.” The procedure
    is not clear and is only described in the experimental part. This
    sentence seems to mean that the matrix W is comprised of
    exemplars (e.g. see
    $http://www.cs.tut.fi/sgn/arg/music/tuomasv/chime-enhancement/pP4_gemmeke.pdf$),
    which is not the case. In supervised NMF framework, the basis matrix
    W is usually pre-trained applying NMF to the training data. It would
    be interesting to compare such approach to the proposed
    K-means procedure.*

     A block diagram has been added to better explain the
    dictionary learning process shown in Figure 7 of the
    updated manuscript. Performing an NMF training have also been tested
    with no significant improvements.

4.  *One type of vehicles (passenger car) is addressed in
    the simulations. Authors may want to comment on why other types like
    motorcycles, which generate a palpable amount of noise and may have
    different acoustic characteristics, are not considered.*

     This important limitation of the presented study is
    now clearly stated as future work of interest in the conclusion :
    ’It has to be noticed that the method would gain to be tested on
    larger dataset with, for instance, motorcycles and other noisy
    traffic elements which are not considered in this study.’

5.  *Is also interesting to know if the distribution of microphones in a
    realistic noise level measurement scenario is taken into account
    during the simulations.*

     The purpose of this study is to propose a tool that
    could be integrated into a sensor network. The distribution of the
    microphone is not a question dealt with in this work. The proposed
    method is designed to perform with a monophonic input. That being
    said, considering the inputs of multiple sensors may help to achieve
    more consistent estimates. This is left for potential future work.

6.  *“The obtained dictionary is expressed with third octave bands”,
    “The spectrogram V, as the dictionary W, is expressed with third
    octave bands (F = 29).” - please, clarify. Perhaps is better to join
    the description of this representation with the section 4.2.1.*

     The manuscript have been updated, see Section 4.2.2:
    ’The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are
    expressed with third octave bands ($F$ = 29). This coarser method
    allows us to reduce the dimensionality and then decrease the
    computation time. Furthermore, by expressing the frequency axis on a
    log frequency axis, the low frequencies, where the traffic energy is
    focused, are described more finely than the high frequencies.
    Experimental validation consistently showed that considering third
    octave bands do not impact the performance of the estimator studied
    in this paper. But, most of all, it is a suited representation to
    this sound environment as this kind of representation is widely used
    in the urban acoustic field, compared to MFCCs for instance.’

7.  *As for the sound level calculation itself, it is not clear if
    A-weighting is applied.*

     As now explicitly stated in Section 4.2.2, ’The
    $A$-weighting of the sound levels is not considered here as it
    decreases the low frequencies levels where the road traffic
    components are mainly present.’

8.  *Also, the terminology is a bit confusing. E.g. Figure 8 “1 second
    equivalent sound pressure level” - to my understanding, the
    equivalent sound pressure level Leq is an average over the specified
    period of time and, therefore, would correspond to a single value.
    If the goal is to eventually simulate the output of a sound pressure
    level meter, some calibration would be needed.*

     The caption of the Figure 8 has been corrected to:
    ’Evolution of 1 second equivalent sound pressure levels’. The
    calibration step is here not necessary as all the sound levels are
    set relatively.

9.  *It is not clear what is the number of iterations that was used for
    each NMF version. It might make sense to use the reduced number of
    iterations for Sup-NMF as it is done for TI-NMF.*

     To better compare the different NMF version, 100
    iterations are now performed for every NMF types, see section 4.2.2.

10. *Is it true that the dictionaries W in Sup-NMF, Ws in Sem-NMF and W0
    in TI-NMF are the same? Authors may want to state it clearly in
    Section 4.2.1.*

     Yes it is, it has been clarified in the updated
    version of the manuscript in Section 4.2.1: ’The 10 versions of the
    built dictionary are then used for NMF. In Sup-NMF, these 10
    versions correspond to $\mathbf{W}$, in Sem-NMF, they correspond to
    the fixed part $\mathbf{W_s}$ and for TI-NMF, they are the initial
    dictionaries $\mathbf{W_0}$ that are next updated.’

Minor points
------------

1.  *“at fixed stations spread all over the cities \[31\] \[30\], which
    would make available of the long-term evolution of the traffic
    noise” - better “which would make the long-term evolution of the
    traffic noise available”.*

     The sentence is rewritten.

2.  *“A two-step process is generally followed :” - there is an extra
    space before the word “followed”*

     Fixed.

3.  *“In the opposite” - better “On the contrary”*

     Fixed.

4.  *“to some extend” should be "to some extent*

     Fixed.

5.  *“by forcing its initiation” should be “by forcing its
    initialization”*

     Fixed.

6.  *Figure 2 caption, please change to “of an urban”, “correspond to”.*

     Fixed.

7.  *Figure 3 caption, the sentence “The 82-nd first elements are
    considered as traffic component.” is confusing.*

     The sentence is rewritten.

8.  *Figure 5 caption, “experience” should be “experiment”?*

     Indeed, fixed.

9.  *“The choice of the dimensions is often made so that F × K + K × N
    $<$ F × N .” - could you please provide a reference for this
    statement?*

     The reference @fevotte_nonnegative_2009 is added.

10. *“must of all” should probably be “most of all”*

     Indeed, fixed.

11. *“The errors are too important for low TIR and this for all
    the sub-classes.” - the sentence is not clear.*

     The sentence is rewritten to: ’Errors are too
    important for all sub-classes at low $TIR$.’